---
title: "Lab 2 - Assignment - C. Thomas - 26Sep2018"
author: "Charlene Thomas"
date: "September 26, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####1. Read Chapters 3, 5, 12, 13 from “R for data science” (online version).

Completed! 

####2. Create a GitHub repository Lab2 (files with R code for this assignment should be added to this repository, do not push csv files for Problem 4 to the GitHub repo).

Completed! 

####3. Write the R code to recreate the graph from the Activity 2 (Problem 4).

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
setwd("~/Documents/BSDS - Fall 2018/Biostatistics I/Data")
activity <- read.csv("lab2.csv")

activity %>%
  gather(`base_sport`, `base_pain`, `base_qol`, 
         `first_sport`, `first_pain`, `first_qol`, `second_sport`, 
         `second_qol`, `second_pain`, key = "time_category", value = "cases") %>%
  separate(time_category, into = c("time", "category")) %>%
  group_by(time, category) %>%
  summarize(mean = mean(cases), sd = sd(cases)) %>%
  ggplot(aes(x = time, y = mean, color =category, group = category)) + 
  geom_point(position = position_dodge(0.08)) +
  geom_line() + 
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.05, position = position_dodge(0.08)) + 
  ylim(0,100) + 
  xlab("Time of report") + 
  ylab("Sample mean and standard deviation")
  
```

#### 4. Download two files:
• coverage.csv - Health Insurance Coverage of the Total Population (2013 - 2016)
• expenditures.csv - Health Care Expenditures by State of Residence (in millions)


#####4.1 Make the data frames “tidy”.
Below in R Code! (Under 4.2) 

#####4.2 Merge two data frames: the resulting data frame should contain information about coverage and expenditures for years 2013-2016. Please note that file expenditures.csv does not contain years 2015-2016.


```{r}
setwd("~/Documents/BSDS - Fall 2018/Biostatistics I/Data")
library(tidyverse)
coverage<- read.csv(file = "coverage.csv", skip =2, check.names = FALSE)
coverage<- coverage[1:52,]

tidy_coverage <- coverage %>%
  gather(`2013__Employer`, `2013__Non-Group`, `2013__Medicaid`, `2013__Medicare`, 
         `2013__Other Public`, `2013__Uninsured`, `2013__Total`, `2014__Employer`, `2014__Non-Group`, `2014__Medicaid`,
         `2014__Medicare`, `2014__Other Public`, `2014__Uninsured`, `2014__Total`, `2015__Employer`, `2015__Non-Group`,
         `2015__Medicaid`, `2015__Medicare`, `2015__Other Public`,`2015__Uninsured`, `2015__Total`, `2016__Employer`,
         `2016__Non-Group`, `2016__Medicaid`, `2016__Medicare`, `2016__Other Public`,`2016__Uninsured`, `2016__Total`,
         key = "year_category", value = "Enrollment (# of People)") %>%
  separate(year_category, into = c("Year", "Coverage Type"))

expenditures<- read.csv(file = "expenditures.csv", skip =2,check.names = FALSE)
expenditures<- expenditures[1:52,]

tidy_expenditures <- expenditures %>% 
  gather(`1991__Total Health Spending`, `1992__Total Health Spending`, `1993__Total Health Spending`,`1994__Total Health Spending`, `1995__Total Health Spending`,
         `1996__Total Health Spending`, `1997__Total Health Spending`, `1998__Total Health Spending`,
         `1999__Total Health Spending`, `2000__Total Health Spending`, `2001__Total Health Spending`,
         `2002__Total Health Spending`, `2003__Total Health Spending`, `2004__Total Health Spending`,
         `2005__Total Health Spending`, `2006__Total Health Spending`, `2007__Total Health Spending`,
         `2008__Total Health Spending`, `2009__Total Health Spending`, `2010__Total Health Spending`,
         `2011__Total Health Spending`, `2012__Total Health Spending`, `2013__Total Health Spending`,
         `2014__Total Health Spending`, key = "year_category", value = "Total Health Spending") %>%
  separate(year_category, into = c("Year", "Type of Spending")) %>%
  select("Location", "Year", "Total Health Spending")

tidy_df <- tidy_coverage %>%
left_join(tidy_expenditures) 
```

####5. Submit a link to the repo “Lab2” via Canvas.

Submitted! 